---
title             : 'Appendices for ""'
shorttitle         : "(Pre-)specifiying effects of interest"

author: 
  - name: "Haiyang Jin"
    affiliation: ""
    address: ""
    email: "haiyang.jin@outlook.com"
    corresponding: yes
    
affiliation:
  - id            : ""
    institution   : ""
  
authornote        : |
  This is the appendices for "".
keywords          : ""
wordcount         : "X"

figurelist        : false
tablelist         : false
footnotelist      : false
linenumbers       : false
mask              : false
draft             : false
floatsintext      : true

documentclass     : "apa6"
classoption       : "man"
numbersections    : false
link-citations    : true

output:
    # papaja::apa6_word:
    #     toc: no
    papaja::apa6_pdf:
        toc: no
        toc_depth: 3
        highlight: default
        latex_engine: xelatex
        
date: "`r format(Sys.time(), '%d-%m-%Y')`"

header-includes   :
- \usepackage{booktabs}
- \usepackage{amsmath}
- \usepackage[american]{babel}
- \usepackage[utf8]{inputenc}
# - \usepackage[T1]{fontenc}
- \usepackage{sectsty} \allsectionsfont{\raggedright} # left-align H1 titles
bibliography: ["`r rbbt::bbt_write_bib('references/references.bib', overwrite = TRUE)`", "references/r-references.bib"]
---

```{r setup, include=FALSE}
## load libraries
library(knitr)
library(tidyverse)
library(lme4)
library(afex)
library(emmeans)
library(papaja)

options(emmeans=list(msg.interaction=FALSE))
theme_set(theme_apa())

# set global chunk options, put figures into folder
options(tinytex.verbose = TRUE)
options(warn=-1, replace.assign=TRUE)
knitr::opts_chunk$set(
  echo = FALSE,
  include = TRUE,
  # warning = FALSE,
  fig.align = "center",
  fig.path = "figures/figure-",
  fig.show = "hold",
  fig.width=7, fig.asp =0.618,
  width = 1800, 
  message = FALSE
)

source(here::here("R", "funcs_sim.R"))

set.seed(2022)
```

\appendix

# (Pre-)specifying effects of interest in practice

## Introduction to the hypothetical study

The aim of this hypothetical training study was to test whether the training protocol is effective in improving participants' ability of recognizing other-race faces. Please refer to the main text for more details. The specific experimental design is as followings:

Independent variables:
- 2 (Group: `train` vs. `control`; between-subject factor)
- 2 (Test: `pre-` vs. `post-test`; within-subject factor)

Dependent variables:
- Behavioral choices (analyzed with sensitivity _d_'; `d`)
- (correct) Response times (analyzed with log-normal transformation; `rt`)

## Data simulation

Let us simulate two data sets for behavioral choices and response times separately with the above design of 30 participants for each group. For simplicity, the average performance for each participant was simulated directly. Please refer to the codes in the open materials for detailed simulation procedures. Table \@ref(tab:simu-true) displays the ground truth for each condition, and the structure of the simulated data is as follows:

```{r}
d_pre <- 2.5
d_delta <- c(0.4, 0.9) # control, train
d_post <- d_pre + d_delta
df_d <- sim_train(d_pre, d_delta, std=0.5, dvname="d")
```

```{r}
str(df_d)
```

```{r simulating rt data}
rt_pre <- 6.4
rt_delta <- c(-0.3, -0.3)
rt_post <- rt_pre + rt_delta 
df_rt <- sim_train(rt_pre, rt_delta, std=0.1, dvname="logrt", logrt=TRUE)
```

```{r}
str(df_rt)
```

```{r simu-true}
tibble(Test = c("pre-test", "post-test"),
       d_control = c(d_pre, d_delta[1]),
       d_train = c(d_pre, d_delta[2]),
       rt_control = c(round(exp(rt_pre)), round(exp(rt_delta[1]))),
       rt_train = c(round(exp(rt_pre)), round(exp(rt_delta[2])))) %>% 
  apa_table(caption='The ground truth for each condition.',
            note='Population parameters used to simulate data for the hypothetical training study.')
```

## Specifying EOI

As discussed in the main text, the EOI for concluding that the training is effective is: ($d^{'}_{post\_train} > d^{'}_{pre\_train}$ and $d^{'}_{post\_train-pre\_train} > d^{'}_{post\_control-pre\_control}$) or ($rt_{post\_train} < rt_{pre\_train}$ and $rt_{post\_train-pre\_train} < rt_{post\_control-pre\_control}$). 




## Data analysis

### Sensitivity d'

Mixed-ANOVA on sensitivity _d_' is first performed based on the 2 $\times$ 2 design with `library(afex)` as follows:


\newpage
