---
title             : "Figures used in the main texts"

author: 
  - name: "Haiyang Jin"
    affiliation: ""
    address: ""
    email: "haiyang.jin@outlook.com"
    corresponding: yes
    
affiliation:
  - id            : ""
    institution   : "New York University Abu Dhabi"
  
authornote        : |
  This is the code for creating the Figures in the main texts.
        
date: "`r format(Sys.time(), '%d-%m-%Y')`"

---

```{r setup, appendix, include=FALSE}
## load libraries
library(knitr)
library(tidyverse)
library(afex)
library(emmeans)
library(ggpubr)
library(here)
library(papaja)

theme_set(theme_apa())

# set global chunk options, put figures into folder
options(warn=-1, replace.assign=TRUE)
knitr::opts_chunk$set(
	echo = FALSE,
	include = TRUE,
	# warning = FALSE,
	fig.align = "center",
	fig.path = "figures/figure-",
	fig.show = "hold",
	fig.width=7, fig.asp =0.618,
	width = 1800, 
	message = FALSE
)

source(here("R", "funcs_sim.R"))
```

# Figure 1

## Figure 1A
The (special) result pattern cannot be used to claim the training effectiveness when the interaction was significant but that simple effect was not.

```{r}
df_1a <- sim_train(c(2.6, 2.4), c(-0.15, 0.15), std=0.5, dvname="d")
```

```{r}
d_aov_1a <- aov_4(d ~ Group * Test + (Test | Subj), df_1a)
(d_emm_1a <- emmeans(d_aov_1a, ~ Group + Test))
```

```{r}
(d_1a_sim <- summary(contrast(d_emm_1a, "revpairwise", by="Group")[2], side = ">", infer=T))
```


```{r}
(d_1a_inter <- summary(contrast(d_emm_1a, interaction="revpairwise"), side = ">", infer=T))
```

```{r}
plot_1a <- as_tibble(d_emm_1a) %>% 
  mutate(Test = fct_recode(Test, `pre-test`="pre", `post-test`="post")) %>% 
  ggplot(aes(Test, emmean, color=Group, group=Group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=Group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15,
                position=position_dodge(.1), show.legend=F) +
  coord_cartesian(ylim = c(0, 4)) +
  labs(y=expression("Sensitivity"~italic("d'")), 
       color="Group", linetype="Group") +
  geom_text(label="textstyle('simple effect:')~italic('p')~'='~.12", x=.5, y = 1, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('interaction:')~italic('p')~'='~.03", x=.5, y = 0.7, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  theme(legend.position=c(.5, .0975),
        legend.direction = "horizontal",
        legend.margin = margin(t = 0, unit='cm'),
        axis.title.x = element_blank()) +
  NULL
```

```{r}
plot_1a
```


## Figure 1B
The result patterns of the simulation data set for _d_' and RT separately.

### d

```{r}
df_d_1b <- sim_train(2.5, c(0.4, 0.9), std=0.5, dvname="d")
```

```{r}
d_aov_1b <- aov_4(d ~ Group * Test + (Test | Subj), df_d_1b)
(d_emm_1b <- emmeans(d_aov_1b, ~ Group + Test))
```

```{r}
(d_1b_sim <- summary(contrast(d_emm_1b, "revpairwise", by="Group")[2], side = ">", infer=T))
```


```{r}
(d_1b_inter <- summary(contrast(d_emm_1b, interaction="revpairwise"), side = ">", infer=T))
```

```{r}
plot_1b_d <- as_tibble(d_emm_1b) %>% 
  mutate(Test = fct_recode(Test, `pre-test`="pre", `post-test`="post")) %>% 
  ggplot(aes(Test, emmean, color=Group, group=Group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=Group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15,
                position=position_dodge(.1), show.legend=F) +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x="Test", y=expression("Sensitivity"~italic("d'")), 
       color="Group", linetype="Group") +
  geom_text(label="textstyle('simple effect:')~italic('p')~'<'~.001", x=.5, y = 1, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('interaction:')~italic('p')~'<'~.001", x=.5, y = 0.7, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  theme(legend.position=c(.5, .0975),
        legend.direction = "horizontal",
        legend.margin = margin(t = 0, unit='cm'),
        axis.title.x = element_blank()) +
  NULL
```

```{r}
plot_1b_d
```

### RT

```{r simulating rt data}
df_rt_1b <- sim_train(6.4, c(-0.1, -0.1), std=0.1, dvname="logrt", logrt=TRUE)
```


```{r}
rt_aov_1b <- aov_4(logrt ~ Group * Test + (Test | Subj), df_rt_1b)
(rt_emm_1b <- emmeans(rt_aov_1b, ~ Group + Test))
```

```{r}
(rt_1b_sim <- summary(contrast(rt_emm_1b, "revpairwise", by="Group")[2], side = "<", infer=T))
```


```{r}
(rt_1b_inter <- summary(contrast(rt_emm_1b, interaction="revpairwise"), side = "<", infer=T))
```

```{r}
plot_1b_rt <- as_tibble(rt_emm_1b) %>% 
  mutate(Test = fct_recode(Test, `pre-test`="pre", `post-test`="post")) %>% 
  mutate(emmean = exp(emmean),
         lower.CL = exp(lower.CL),
         upper.CL = exp(upper.CL)) %>% 
  ggplot(aes(Test, emmean, color=Group, group=Group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=Group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15,
                position=position_dodge(.1), show.legend=F) +
  coord_cartesian(ylim = c(450, 650)) +
  labs(x="Test", y="Correct response times (ms)", 
       color="Group", linetype="Group") +
  geom_text(label="textstyle('simple effect:')~italic('p')~'<'~.001", x=.5, y = 500, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('interaction:')~italic('p')~ '=' ~.32", x=.5, y = 485, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  theme(legend.position="none",
        axis.title.x = element_blank()) +
  NULL
```

```{r}
plot_1b_rt
```

### Combination
```{r}
plot_1b <- ggarrange(plot_1b_d, plot_1b_rt)
plot_1b
```


## Figure 1C
The three scenarios where both that simple effect and interaction were significant.

### Scenario 1

```{r}
df_1c1 <- sim_train(2.5, c(0.25, 0.6), std=0.5, dvname="d")
```

```{r}
d_aov_1c1 <- aov_4(d ~ Group * Test + (Test | Subj), df_1c1)
(d_emm_1c1 <- emmeans(d_aov_1c1, ~ Group + Test))
```

```{r}
(d_1c1_sim <- summary(contrast(d_emm_1c1, "revpairwise", by="Group"), side = "=", infer=T))
```


```{r}
(d_1c1_inter <- summary(contrast(d_emm_1c1, interaction="revpairwise"), side = "=", infer=T))
```

```{r}
plot_1c1 <- as_tibble(d_emm_1c1) %>% 
  mutate(Test = fct_recode(Test, `pre-test`="pre", `post-test`="post")) %>% 
  ggplot(aes(Test, emmean, color=Group, group=Group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=Group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15,
                position=position_dodge(.1), show.legend=F) +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x="Test", y=expression("Sensitivity"~italic("d'")), 
       color="Group", linetype="Group") +
  geom_text(label="simple~effect[' control']:italic('p')~'='~.005", x=.5, y = 1.3, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('simple effect')[' train']:~italic('p')~'<'~.001", x=.5, y = 1, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('interaction:')~italic('p')~'=' ~.023", x=.5, y = 0.7, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  theme(legend.position="none", #c(.5, .0975),
        legend.direction = "horizontal",
        legend.margin = margin(t = 0, unit='cm'),
        axis.title.x = element_blank()) +
  NULL
```

```{r}
plot_1c1
```


### Scenario 2

```{r}
df_1c2 <- sim_train(2.5, c(0, 0.6), std=0.5, dvname="d")
```

```{r}
d_aov_1c2 <- aov_4(d ~ Group * Test + (Test | Subj), df_1c2)
(d_emm_1c2 <- emmeans(d_aov_1c2, ~ Group + Test))
```

```{r}
(d_1c2_sim <- summary(contrast(d_emm_1c2, "revpairwise", by="Group"), side = "=", infer=T))
```


```{r}
(d_1c2_inter <- summary(contrast(d_emm_1c2, interaction="revpairwise"), side = "=", infer=T))
```

```{r}
plot_1c2 <- as_tibble(d_emm_1c2) %>% 
  mutate(Test = fct_recode(Test, `pre-test`="pre", `post-test`="post")) %>% 
  ggplot(aes(Test, emmean, color=Group, group=Group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=Group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15,
                position=position_dodge(.1), show.legend=F) +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x="Test", y=expression("Sensitivity"~italic("d'")), 
       color="Group", linetype="Group") +
  geom_text(label="textstyle('simple effect')[' control']:italic('p')~'='~textstyle('.90')", x=.5, y = 1.3, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('simple effect')[' train']:~italic('p')~'<'~.001", x=.5, y = 1, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('interaction:')~italic('p')~'<' ~.001", x=.5, y = 0.7, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  theme(legend.position = c(.5, .0975),
        legend.direction = "horizontal",
        legend.margin = margin(t = 0, unit='cm'),
        axis.title.x = element_blank()) +
  NULL
```

```{r}
plot_1c2
```


### Scenario 3

```{r}
df_1c3 <- sim_train(2.5, c(-0.2, 0.6), std=0.5, dvname="d")
```

```{r}
d_aov_1c3 <- aov_4(d ~ Group * Test + (Test | Subj), df_1c3)
(d_emm_1c3 <- emmeans(d_aov_1c3, ~ Group + Test))
```

```{r}
(d_1c3_sim <- summary(contrast(d_emm_1c3, "revpairwise", by="Group"), side = "=", infer=T))
```


```{r}
(d_1c3_inter <- summary(contrast(d_emm_1c3, interaction="revpairwise"), side = "=", infer=T))
```

```{r}
plot_1c3 <- as_tibble(d_emm_1c3) %>% 
  mutate(Test = fct_recode(Test, `pre-test`="pre", `post-test`="post")) %>% 
  ggplot(aes(Test, emmean, color=Group, group=Group)) +
  geom_point(size = 2, position=position_dodge(.1)) +
  geom_line(aes(linetype=Group), size = 1, position=position_dodge(.1)) +
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.15,
                position=position_dodge(.1), show.legend=F) +
  coord_cartesian(ylim = c(0, 4)) +
  labs(x="Test", y=expression("Sensitivity"~italic("d'")), 
       color="Group", linetype="Group") +
  geom_text(label="textstyle('simple effect')[' control']:italic('p')~'='~.037", x=.5, y = 1.3, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('simple effect')[' train']:~italic('p')~'<'~.001", x=.5, y = 1, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  geom_text(label="textstyle('interaction:')~italic('p')~'<' ~.001", x=.5, y = 0.7, # 
            color="black", hjust=0, show.legend = F, parse = T) +
  theme(legend.position="none", #c(.5, .0975),
        legend.direction = "horizontal",
        legend.margin = margin(t = 0, unit='cm'),
        axis.title.x = element_blank()) +
  NULL
```

```{r}
plot_1c3
```

### Combination
```{r}
plot_1c <- ggarrange(plot_1c1, plot_1c2, plot_1c3,
                     nrow = 1)
plot_1c
```


## Combined figure
```{r fig.width=4.5, fig.height=3.2}
plot1<- ggarrange(ggarrange(plot_1a, plot_1b,
          labels = c("A", "B"),
          ncol = 2,
          widths = c(1,2)),
          plot_1c, 
          nrow = 2, labels = c("", "C"),
          common.legend = T)
plot1
```


```{r}
ggsave("figure1.png", plot1, width=9.5, height = 6.4)
```



# Session information
```{r}
sessionInfo()
```


```{r}
# ggsave(file=here::here("images", "inter_simple_sig.png", width = 7.2, height = 5.4)
```

